Agents and you, a brief summary

### 2015-11-02 agents/summery.tex
added a bunch of information, halfway the last slide of markov

>>>>

\subsection{Example Markov reward process}
Consider the following markov reard process. Suppose a discount factor
$\gamma = 0.5$.

\[ \left\{ \begin{matrix}
		s: & s(r=1, p=0.4) & ; & a(r=4, p=0.6), \\
		a: & s(r=3, p=0.3) & ; & a(r=8, p=0.7), \\
		b: & s(r=2, p=0.5) & ; & b(r=5, p=0.5). \\
	\end{matrix}
\right. \]

\subsubsection{Transition diagram}
\begin{tikzpicture}[->, >=stealth', auto, semithick, node distance=3cm]
\tikzstyle{every state}=[fill=white,draw=black,thick,text=black,scale=1]
\node[state]    (S)                     {s};
\node[state]    (A)[right of=S]   {a};
\node[state]    (B)[left of=S]   {b};
\path
(S) edge[loop above]     	node{p=0.4, r=1}(S)
    edge[above]     		node{p=0.6, r=4}(A)
(A) edge[bend left,below]   node{p=0.3, r=3}(S)
	edge[loop above]     	node{p=0.7, r=8}(A)
(B) edge[above]    			node{p=0.5, r=2}(S)
	edge[loop above]     	node{p=0.5, r=5}(B);
\end{tikzpicture}
\subsubsection{Matrici}
\paragraph{Probability transition matrix}
\[
	P =
\begin{pmatrix}
	0.4	& 0.6 & 0 	& \\
	0.3	& 0.7 & 0	& \\
	0.5	& 0 & 0.5	& \\
\end{pmatrix}
\]
\paragraph{Immediate reward matrix}
\[
	R =
\begin{pmatrix}
	1	& 4 & 0 	& \\
	3	& 8	& 3	& \\
	2	& 0 & 5	& \\
\end{pmatrix}
\]
\subsubsection{Value evalutaion}
apply:
\[V(i)=\sum_{all\mbox{\ }j} p_{ij}[r_{ij} + V(j)]\]

\[ \left\{ \begin{matrix}
		V(s) = 0.4(1+\gamma V(s)) + 0.6(4+\gamma V(a)), \\
		V(a) = 0.3(3+\gamma V(s)) + 0.7(8+\gamma V(a)), \\
		V(b) = 0.5(2+\gamma V(s)) + 0.5(5+\gamma V(b)).
	\end{matrix}
\right. \]

With abusing notation:

\[ \left\{ \begin{matrix}
		s = 0.4(1+\gamma s) + 0.6(4+\gamma a), \\
		a = 0.3(3+\gamma s) + 0.7(8+\gamma a), \\
		b = 0.5(2+\gamma s) + 0.5(5+\gamma b).
	\end{matrix}
\right. \]

Just fill in the values recursevily to perform value iteration.

\subsection{Decision processes}
I'm gonna cover this subject poorly because its dead simple coming from an
IT background.

\begin{tabular}{ll}
	word & definition \\ \toprule
	decision tree & a simple directed tree stucture \\
	states & represented by verticies.\\
	Immediate rewards & represented by numbers along the edges. \\
	Decision alternatives & the outgoing edges, also called possible actions. \\
	Decision point & A state where a decision can be made. \\
	Finite decision tree & \brcell{A tree with a finite number of states, \\
		{\footnotesize Makes you wonder if there could be infinite number 
		of decisions in such a tree.}} \\
\end{tabular}
\\
The set of possible actions at state $i \in I$ is denoted by $A_i$.
The immediate reward for action $j$ at $i$ is denoted by $r_{ij}$.

\subsubsection{Policy}
A \emph{policy} specifies for every decision point $s$ a probability
distribution over $A_s$. A policy is typically denoted by $\pi, \tau\dots$.
If at $s$ the probability distribution puts all weight on one choice, the
policy is said to be \emph{pure} at $s$. Otherwise the policy is said to
be \emph{mixed} at $s$.

A \emph{run} of a policy is the \emph{actual execution} of that policy.
An \emph{realisation}, or \emph{sample path} of a policy is the trace of
that run.

The \emph{payoff} or \emph{total reward} of a run is the total sum of rewards
acuumulated in its realisation.

\subsubsection{Value function}
A \emph{value function} is any function $V: S \to \mathbb{R}$. The
\emph{expected payoff of a policy $\pi$}, given a certain starting state $s$:
	\[E[\pi|s_0=s]\]
	is called the value of $s$ under $\pi$, and is denoted by $V^\pi(s)$.
Value functions are compared in the following way:

\[ V_1 \le V_2 \Leftrightarrow_{Def} V_1(s) \le V_2(s), \mbox{for all states } s.\]

Policy $\pi$ is \emph{as good as} policy $\tau$ if $V^\tau \le V^\pi$. A policy
is \emph{optimal} if it is as good as any other policy. A value function is
\emph{optimal} if it is induced by a policy, and is as large as any other
such value function.

\subsubsection{Policy = strategy}
% this is the first what came to mind when I saw the policies.
In game theory, a policy is called a \emph{strategy} and a pure strategy is
sometimes called \emph{stationary} or \emph{deterministic}. A mixed strategy
is sometimes called \emph{probabilistic, non-stationary,} or
\emph{non-deterministic}.
% Imaginative with their non-*

An example of a pure policy is
\[ \tau = \{(r,b),(a,d),(b,e),(d,h),(f,i)\} \]

All the decisions are made, even if you already cut them off.
So you say (Start, End) for each node with choice.

Mixed policy:
\[ \tau = \{(r,\{(a, 0.2), (b,0.8)\}),(a,d),(b,\{(e, 0.7), (f,0.3)\}),
(d,h),(f,i)\} \]
% yes some really poor langauge design over here.

\subsubsection{Expected payoff}
$\pi$'s expected payoff, $E[\pi|s_0=r]$, can be computed recursivly with the
law of total expectation.

\subsubsection{Backward induction}
A decision between alternatives that leads to certain outcomes, is a certainty
itself.

Backward induction is applicable in \emph{directed acyclic graphs}(DAGs).

\subsubsection{Decision problem}
A \emph{decision problem} amounts to determining the set $\Pi^*$, where:

\[ \Pi = max_{\pi \in \Pi}(payoff(\pi))\]

Where $\Pi$ is the set of all policies.

For finite decision trees: Supppose at least one optimal policy exists.
If this policy is mixed, it may be replaced by an optimal pure policy,
such that, at every state, the payoff of the pure policy equals the expected
payoff of the mixed policy.

\subsubsection{Problems}
\paragraph{infinite width} no best policy, because you're valuating forever.
\paragraph{infinite depth} How to compare infinite total rewards (can't).
\paragraph{discount factor} The right decision is very dependent on the
discount factor. Counter intiuitive choices might be better for example.

\subsection{Utility pumps}
If there is no discount factor policy is irrelivant (the reward will be
infinite no matter what you do). A tool is needed to kill infinity.
For example a finiti horizon, discount factor, or long-run average
payoffs.

In this course we will just use discount factors.
